---
title: "Untitled"
author: "xinyi zheng"
date: "2023-11-23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 2

## load necessary packages
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(broom)
library(rnoaa)
library(modelr)
```

## import the data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## Function to perform linear regression and calculate required statistics
```{r}
analyze_bootstrap_sample <- function(sample) {
  model <- lm(tmax ~ tmin + prcp, data = sample)
  r_squared <- glance(model)$r.squared
  coefs <- coef(model)
  log_b1_b2 <- log(abs(coefs["tmin"] * coefs["prcp"])) # using abs() to avoid NaNs from negative values
  return(c(r_squared, log_b1_b2))
}

data_results <- analyze_bootstrap_sample(weather_df)
```

## Bootstrap analysis
```{r}
set.seed(1)
bootstrap_results <- replicate(5000, {
  sample_indices <- sample(1:nrow(weather_df), replace = TRUE)
  sample_data <- weather_df[sample_indices, ]
  analyze_bootstrap_sample(sample_data)
}, simplify = "array")
```

## Convert results to a data frame
```{r}
bootstrap_results_df <- data.frame(t(bootstrap_results))
names(bootstrap_results_df) <- c("r_squared", "log_b1_b2")
```

## Remove NA and NaN values
```{r}
bootstrap_results_df <- na.omit(bootstrap_results_df)
```

## Plotting distributions
```{r}
ggplot(bootstrap_results_df, aes(x = r_squared)) +
  geom_histogram(binwidth = 0.01, fill = "blue", color = "black") +
  labs(title = "Distribution of r^2")

ggplot(bootstrap_results_df, aes(x = log_b1_b2)) +
  geom_histogram(binwidth = 0.1, fill = "red", color = "black") +
  labs(title = "Distribution of log(β̂1 * β̂2)")
```

## Calculating confidence intervals with na.rm = TRUE
```{r}
quantile(bootstrap_results_df$r_squared, c(0.025, 0.975), na.rm = TRUE)
quantile(bootstrap_results_df$log_b1_b2, c(0.025, 0.975), na.rm = TRUE)

```

## Findings
note:When I calculated log(beta_1 * beta_2), I used the function abs() to avoid NaNs from negative values.

From the distribution plot, we find that r^2 approximately follows normal distribution.( need to check) The mean estimate is 0.92 with 95% CI (XXX,XXX). 

The distribution of log (beta_1 * beta_2) is skewed to the left, so it doesn't follow normal distribution for sure. So we need to use bootstrap method to estimate the 95% confidence interval. The mean estimate is -6.46 with 95% CI (XXX, XXX). 

 
# Problem 3

## Import dataset
```{r}
birth_data <- read.csv('data/birthweight.csv')
```

## Convert categorical variables to factors
```{r}
categorical_vars <- c('babysex', 'frace', 'mrace', 'malform')
birth_data[categorical_vars] <- lapply(birth_data[categorical_vars], factor)

# Check for missing data
summary(birth_data)
# by viewing summary results, there is no missing data in this file. 

```

## Propose a regression Model based on the results of stepwise selection. So it is a data-driven approach.
```{r}
full.model <- lm(bwt ~., data = birth_data)
```

## Stepwise regression model
```{r}
step.model <- MASS::stepAIC(full.model, direction = "both", 
                      trace = FALSE)
summary(step.model)

my_model <- lm(bwt ~ babysex+bhead+blength+delwt+fincome+gaweeks+mheight+mrace+parity+ppwt+smoken, data = birth_data)


```

## Base on my hypothesis
```{r}
# Check residuals
birth_data %>% 
  add_predictions(my_model) %>% 
  add_residuals(my_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Fitted values", y = "Residuals")

# Model 1: Length and Gestational Age
model_1 <- lm(bwt ~ blength + gaweeks, data = birth_data)

# Model 2: Head Circumference, Length, Sex, and Interactions
model_2 <- lm(bwt ~ bhead * blength * babysex, data = birth_data)

# Cross-Validation
cv_models <- birth_data %>% crossv_mc(50) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    my_model  = map(train, \(df) lm(bwt ~ babysex+bhead+blength+delwt+fincome+gaweeks+mheight+mrace+parity+ppwt+smoken, data = df)),
    model_1  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model_2  = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data=df))) %>% 
  mutate(
    rmse_mymodel = map2_dbl(my_model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model_1 = map2_dbl(model_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model_2 = map2_dbl(model_2, test, \(mod, df) rmse(model = mod, data = df))
    )

cv_models |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()


```

